\section{Approach to get to 80\% milestone}
The 80\% milestone is a background research report. Based on the deadline (Nov 10, 2017), we divide our process into 3 phases.

\subsection{Phase 1: Get familiar with ABC notation}
\textbf{Time} Oct 27, 2017 --- Oct 31, 2017

\noindent Familiarize ourselves with ABC notations features. Read through the abc notation website. Download the software related to ABC and play with it, for example, easyABC. Each team member will write 2 small music pieces(can be very short) using the existing ABC notation, submit them to the git repo, and demo them on the project meeting on Oct 30. Discuss our own music-writing experience and technical difficulties with ABC and document our conclusions in a google doc.

\subsection{Phase 2: Read the full ABC language specification}
\textbf{Time} Nov 1, 2017 --- Nov 5, 2017

\noindent Read through the full ABC language specification, record our thoughts and discoveries in a google doc. Discuss online or in person to decide which parts can be used directly or after some modification in our ABCD language, and document these in our draft of ABCD specification. Record in google docs why such choices have been made. Estimate what we need to do for ABCD. The result of this phase will be a pdf file that is part of our ABCD language specification, which lists all the borrowed notations from ABC.

\subsection{Phase 3: Criticize ABC and develop a draft ABCD}
\textbf{Time} Nov 6, 2017 --- Nov 9, 2017

\noindent Read discussions on stackexchange and blogs to see people's opinion and suggestion on ABC. List ABC's advantages and disadvantages and make clear the value of developing ABCD. Integrate our conclusions in phase 1 and phase 2. Discussion our potential project. Go through the citations. The result of this phase will be our background research report.

